---
title: "hiearchical clustering"
description: |
  A featured code project.
author:
  - name: Haley Grant
    url: {}
date: 03-14-2021
output:
  distill::distill_article:
    code_folding: true
    self_contained: false
---
For this project I created a dendrogram using hierarchical clustering by complete linkage. The figure demonstrates the relationship between sites based on water chemistry.[^1] Hierarchical clustering means the sites are grouped into clusters based on similarity, while complete linkage means the clusters are connected by smallest maximum distance. Dendrograms visualize this organizational structure. To make my dendrogram I used `ggdendrogram`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
library(naniar)
```

```{r}
sbc_lter <- read_csv(here("sbc_lter_registered_stream_chemistry.csv")) %>% 
  replace_with_na(replace = list(nh4_uM = -999,
                                 no3_uM = -999,
                                 po4_uM = -999,
                                 tdn_uM = -999,
                                 tdp_uM = -999,
                                 tpc_uM = -999,
                                 tpn_uM = -999,
                                 tpp_uM = -999,
                                 tss_mgperLiter = -999,
                                 spec_cond_uSpercm = -999)) %>% 
  select(!timestamp_local) %>% 
  pivot_longer(cols = !site_code, names_to = "pollutant")

site_summary <- sbc_lter %>% 
  group_by(site_code) %>% 
  mutate(mean = mean(value, na.rm = TRUE)) %>% 
  na.omit() %>% 
  select(site_code, mean) %>% 
  unique()

sbc_lter_scaled <- site_summary %>%
  ungroup() %>% 
  select(mean) %>% 
  scale()

rownames(sbc_lter_scaled) <- site_summary$site_code

euc_distance <- dist(sbc_lter_scaled, method = "euclidean")

hc_complete <- hclust(euc_distance, method = "complete")

ggdendrogram(hc_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(title = "Hiearchical multivariate clustering by complete linkage",
       x = "Site Code",
       y = "")
```

[^1]: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174.
